---
title: "h9"
author: "grienne"
date: "November 14, 2018"
output: html_document
---

```{r Base}

library(xlsx)
library(dplyr)
library(ggplot2)
library(readxl)
library(reshape2)


options(scipen=999)
dat <- read_excel("C:/Users/Custom PC/Desktop/PhD/CPH_738_Olga/h9/Arit+Harvanko+EC+Survey+data.xlsx")
names(dat)



## subset to only those variables we are going to work with
d <- dat[,c(3:5, 54:61, 70:71)]
names(d) <- sub(".*\\ ", "", names(d))
names(d)[1:3] <- c("age", "sex", "edu")
names(d)[12:13] <- c("VG", "PG")

d$PG.level <- as.factor(with(d,
                             ifelse(PG < 34, "1",
                             ifelse(PG >= 34 & PG < 67, "2",
                                    "3"))))
d$VG.level <- as.factor(with(d,
                             ifelse(VG < 34, "1",
                             ifelse(VG >= 34 & VG < 67, "2",
                                    "3"))))
summary(d)  



tmp.dat <- melt(d, id.vars = c("age", "sex", "edu", "VG", "PG", "PG.level", "VG.level"),
                variable.name = "effect", value.name = "score")
tmp.dat <- na.omit(tmp.dat)

## ---------------------------------------------------
## Relationship with PG.level
## everyone
p = ggplot(aes(x = PG.level, y = score, fill = effect), data = tmp.dat)
p + geom_boxplot() +
    geom_point(alpha = 0.3, size = 3) +
    facet_grid(. ~ effect) +
    theme_bw()


```

Question 1 

1. To explore the association between dizziness and PG/VG ratio, would you use the
ordered PG-levels or the ordered VG-levels. Why or why not?
  PG levels, the article suggested that greater PG level proportions were associated with increased nicotine concentration, which by hypothesis would create more dizziness likelihood. The graph shows that dizziness increases with score at PG levels. 

```{r Question 2}

require(foreign)
require(ggplot2)
require(MASS)
require(Hmisc)
require(reshape2)

attach(d)
detach(d)

tmp.dat <- melt(d, id.vars = c("age", "sex", "edu", "VG", "PG", "PG.level", "VG.level"),
                variable.name = "effect", value.name = "score")

##how does it know which column to use for effect?

tmp.dat <- na.omit(tmp.dat)

p1 = ggplot(aes(x = PG.level, y = score, fill = effect), data = tmp.dat)
p1 + geom_boxplot() +
    geom_point(alpha = 0.3, size = 3) +
    facet_grid(. ~ effect) +
    theme_bw()

## males
pm = ggplot(aes(x = PG.level, y = score, fill = effect), data = subset(tmp.dat, sex == 0))
pm + geom_boxplot() +
    geom_point(alpha = 0.3, size = 3) +
    facet_grid(. ~ effect) +
    theme_bw()


## females
pf = ggplot(aes(x = PG.level, y = score, fill = effect), data = subset(tmp.dat, sex == 1))
pf + geom_boxplot() +
    geom_point(alpha = 0.3, size = 3) +
    facet_grid(. ~ effect) +
    theme_bw()


## edu - some
psome = ggplot(aes(x = PG.level, y = score, fill = effect), data = subset(tmp.dat, edu == "some college"))
psome + geom_boxplot() +
    geom_point(alpha = 0.3, size = 3) +
    facet_grid(. ~ effect) +
    theme_bw()


## females
pbs = ggplot(aes(x = PG.level, y = score, fill = effect), data = subset(tmp.dat, edu == "bachelors degree"))
pbs + geom_boxplot() +
    geom_point(alpha = 0.3, size = 3) +
    facet_grid(. ~ effect) +
    theme_bw()



q2_fit_1 <- polr(PG.level ~ dizziness * sex, data = d, Hess = TRUE)

```

Question 3

Those who had no female were less likely to switch if they reported dizziness


```{R Question 3}

summary(q2_fit_1)
ci <- confint(q2_fit_1)
exp(cbind(OR = coef(q2_fit_1), ci))


#ask why the coefficients arent coming out
```

Question 4

```{r Question 4}


## two cuts
c12 <- q2_fit_1$zeta[1]
c23 <- q2_fit_1$zeta[2]

## visualize the fitted model
invlogit <- function(x) 1 / (1 + exp(-x))
expected <- function (x, c1.5, c2.5, sigma){
    p1.5 <- invlogit ((x-c1.5)/sigma)
    p2.5 <- invlogit ((x-c2.5)/sigma)
    return ((1*(1-p1.5) + 2*(p1.5-p2.5) + 3*p2.5))
}

## Females: y = -3.34 + (-0.01027 + 0.03998) * taste
## align cuts with taste for males
b0 <- coefficients(q2_fit_1)[2]
b1 <- coefficients(q2_fit_1)[1] + coefficients(q2_fit_1)[3]

(c1.5 <- (c12 - b0)/ b1)
(c2.5 <- (c23 - b0)/ b1)
(sgm <- 1 / b1)

p = ggplot()
p + geom_point(aes(x = dizziness, y = PG.level), data = na.omit(subset(d, sex == 1)),
               size = 3, alpha = 0.4) +
    
    geom_segment(aes(x = c1.5, y = 1, xend = c1.5, yend = 2)) +
    geom_segment(aes(x = c2.5, y = 2, xend = c2.5, yend = 3)) +
    theme_bw()

```
